Kids SUMs Summary

### 2020-01-24 sums_data_cleaning.Rmd
new functions for SUMs data summary

thresholds <- sums_clean %>% 
  group_by(area, home, winter_id) %>%
  mutate(temp_sums_40 = if_else(temp_c_sums >= 40, 1, 0),
         temp_sums_40_event = if_else(temp_sums_40 == 1 & lag(temp_sums_40) == 0, "Start", "Null"),
         temp_sums_40_event = if_else(is.na(temp_sums_40_event), "Start", temp_sums_40_event),
         temp_sums_40_event = if_else(temp_sums_40 == 1 & temp_sums_40_event == "Null",
                                      "Event", temp_sums_40_event),
         temp_sums_40_event = if_else(temp_sums_40_event == "Event" &
                                      lead(temp_sums_40_event) == "Null" &
                                      lead(temp_sums_40_event, 2) == "Null" &
                                      lead(temp_sums_40_event, 3) == "Null", "End", temp_sums_40_event)) %>% 
  group_by(area, home, winter_id, date_sums, temp_sums_40_event) %>% 
  summarize(n = n()) %>% 
  group_by(temp_sums_40_event) %>% 
  summarize(mean(n))

table(thresholds$temp_sums_40_event)
```

>>>>

thresholds_test <- sums_clean %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= 1.5, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= 1.5 & lead(next_event) <= 1.5, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds_test

# Function for calculating heating events using specified threshold temps
# Also specify the "gap" time between events
# Try another function for % day above threshold
thresholds_function <- function(data, temp, gap) {

thresholds <- data %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= gap, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds

}

thresholds_function(sums_clean, 40, 2)

# Function to look at percentage of time spent above a threshold
thresholds_percent <- function(data, temp) {

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

}

thresholds_percent(sums_clean, 40)
```

### 2020-01-24 sums_summary.Rmd
new functions for SUMs data summary

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 23 Jan 2020"
output: pdf_document
---

@@

>>>>

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 24 Jan 2020"
output: pdf_document
---

@@

# Function for calculating heating events using specified threshold temps
## Also specify the "gap" time between events
```{r}
thresholds_function <- function(data, temp, gap) {

thresholds <- data %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= gap, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds

}

thresholds_function(sums_clean, 40, 2)
```

# Function to look at percentage of time spent above a threshold
```{r}
thresholds_percent <- function(data, temp) {

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

}

thresholds_percent(sums_clean, 40)
```
